The Anatomy of a Temporal Summarization System

### 2015-02-23 methods.tex
Added experiments and results

\section{Methods}
\label{sec:methods}
\subsection{Data}\label{sec:data}

Our primary datasets are the TREC 2013/2014 Stream Corpus, 
a ?tb corpus of hourly 
web crawls from October 2011
through mid February 2013
\cite{frank2012building}.\footnote{\url{http://trec-kba.org/kba-stream-corpus-2014.shtml}}
%KM - Keep neutral from Trec
%As per the TS specifications,
All summary sentences come from processing this
corpus in a time aligned manner.
%KM - Are you using both 2013 and 2014 corpora.
%KM - Mentioning sentence ids at this point is too low level.
%KM - I would probably just drop this sentence. since you say above that you
%use both.
%While the 2014 corpus encompasses the 
%2013 corpus, we use the 2013 corpus for evaluating events from the 2013 track
%year because sentence ID's used in evaluation changed from year to year.

%KM - As far as I can see this is the first time you've mentioned this.
%KM - This paragraph is not clearly understandable. Is this the training data? 
For our semantic similarity component (\cref{subsec:semsim}),
domain specific term-sentence matrices were built using subsections of
Wikipedia. For each event type we identified one or more related Wikipedia
categories and collected all pages that were listed under that category.
For all pages collected, we used the latest possible revision date before the
start of the Stream Corpus.

We use the same Wikipedia corpora to construct the language models used
in our salience prediction model (\cref{subsubsec:lm}).

\begin{figure*}
        \begin{center}
\begin{tabular}{| p{5cm} c c |}
\hline
Event & WP Categories & No. Docs/Sents/Words\\
\hline \hline
Boston\_Marathon\_bombings \newline
Christopher\_Dorner\_shootings\_and\_manhunt \newline 
 In\_Amenas\_hostage\_crisis
& Terrorism, Mass Shootings & 33,732/1,139,588/26,201,659  \\
\hline
 Cyclone\_Oswald \newline
Early\_2012\_European\_cold\_wave \newline
February\_2013\_noreaster \newline & Weather events & 35,554/591,850/12,794,438  \\
\hline
Costa\_Concordia\_disaster & Accidents & 22,874/732,945/16,520,242 \\
\hline
Chelyabinsk\_meteor & Earthquakes & 14,515/283,509/6,135,803  \\
\hline
Port\_Said\_Stadium\_riot \newline
2012\_Afghanistan\_Quran\_burning\_protests \newline
2011-13\_Russian\_protests \newline
2012\_Romanian\_protests \newline
2012-13\_Egyptian\_protests \newline
2013\_Bulgarian\_protests\_against\_the\_Borisov\_cabinet \newline
2013\_Shahbag\_protests & Activism\_by\_type & 464,657/11,254,122/250,172,896  \\
\hline
\end{tabular}
\caption{Domain specific Wikipedia corpora for semantic similarity and 
language models. }
\end{center}
\end{figure*}

\subsubsection{Events}

%KM - You said this already so I think you can make it simpler. 
%We use the event set developed for the TREC Temporal Summarization track 
%for years 2013 and 2014.
The TREC 2013 and 2014 Stream corpus contains  10 and 15 events respectively. We omit
one event from the 2013 track because there were no news documents in the 
corpus during the duration of the event, yielding a final event set of 24 
events. Meta-data provided for each event includes the name of the event
as it appears on Wikipedia, a short text query for the event, 
the type of event, and the 
duration of the event
%KM - the below parens is not clear. is ``to monitor'' missing from the above
%sentence? I think get rid of the parens below.
%(the duration for the system 
%to monitor is not necessarily the complete timespan of the event).
%KM - Let's get rid of all references to ``the TS system'' If you want to have
%a system name, choose a better one. 
All data except the Wikipedia name is exposed to
systems that participate in TREC.
%the TS system to use.

\subsubsection{Nuggets}
For evaluation purposes, important pieces of information, hereafter 
referred to as nuggets, were extracted from each event's Wikipedia page.
%KM - What assessors? Was this done by TREC? Make clear. I added ``TREC'' after
%reading ahead.
%KM - What do you mean ``graded''? Do you mean they graded your nuggets? Or do
%you mean they ``ranked'' the nuggets by importance. Please make clear.
TREC assessors graded nuggets by importance, and were also able to timestamp
%KM - wrong wording. I've changed.
%their
%existance
when the event represented by each nugget occurred 
using the timestamp of the page's revision history. Nuggets are 
usually a short to medium length clause containing an atomic piece of
information.

%KM - I changed title to make clearer
\subsubsection{Gold Standard Matches}
TREC assessors provided judgements on a subset of participant updates each year
of the track. Judged updates are matched to one or more nuggets (a real 
sentence often contains more than one atomic piece of information) or marked
not relevant.

%KM - I don't see why overlap in particpant update sets would give you low
%coverage. In fact, it might you give you broader coverage than if all
%participants extracted the same sentences.  I removed first portion of the
%sentence, but you can put back and clarify if you think I interpreted
%incorrectly. 
%There was not much overlap in the participant update sets so
Overall coverage of
%KM - Shouldn't this be nuggets?
relevant sentences
relevant nuggets in the corpus  by participant systems' generated sentences
is small (percent?).
%KM - This is complicated. More explanation is needed.
%KM - The sentence below sounds like how you evaluate your system and not like
%the judgment corpus.
%KM not sure this owuld go here. Shouldn't it go in evaluation? Here you would
%just talk about the corpus
To evaluate new system output on the same event, we compare it against the
corpus of matched sentences. Comparing it against the nuggets would not yield
meaningful results as the nuggets are much smaller than the sentences of the
corpus.
Because there are so few matched sentences, we score all new system updates as positive
%To expand the judgement corpora, we include all system updates as positive
nugget matches if they are $< .2$ close in Levenshtein distance to a human
assessed matched update.

%KM - Got this far.
\subsection{Experiment Design}

\subsubsection{Salience Models}
For each event (both in the complete system and feature ablation experiments)
we randomly sample 1000 sentences after relevant document filtering and 
over the duration of the event 
(as determined by 
TREC; avg. run length is ? days) of the event. Nugget similarities
are computed for each sentence and are standardized with respect to 
the sample mean and variance. After standardization, the maximum nugget similarities are computed for each sentence. 
We fit a Gaussian process for each event, using the maximum similarities 
as our regression target, yielding 24 models in total. Kernel lengthscale
parameters are also individually fit for each model.
When making event-specific predictions, we take the average prediction of the
23 other models.

\subsubsection{Estimating Language Models}
We use two trigram language models, trained using the SRILM toolkit \cite{stolcke2002srilm}, taking as features the average log probability (i.e. the sentence's total log probability normalized by sentence length) from each model.  This first model is trained on 4 years (2005-2009) of articles from the Gigaword corpus.  Specifically, we use articles from the Associated Press and the New York Times. The second model is a domain specific language model. We build a corpus of Wikipedia articles for each event type, consisting of documents from a related Wikipedia category. E.g. for earthquakes, we collect pages under the category \emph{Category:Earthquakes}. This model assigns higher probability to sentences that are focused on the given domain.

% For each event, we retrieve all sentences using the document retrieval and content selection steps outlined in the previous selection. We then sample with replacement 200 sentences from this collection, and extract non-lexical features for each sentence to construct our design matrix $\mathbf{X}$. To build our response variables $\mathbf{y}$ we compute the maximum semantic similarity of each input sentence to the gold nugget sentences. We fit a $\operatorname{GP}$ with an RBF kernel to this data, optimizing kernel parameters with the scaled  conjugate gradient method. This sampling procedure is repeated 100 times for each of the ? 2013 TS events, yielding \fdcomment{?} total models. In our salience predictions for the 2014 events, we take the mean prediction of all models.

\subsubsection{Redundancy Threshold}

For every time interval with input, a clustering algorithm is
guarranteed to produce at least one output; 
this can significantly hurt the performance of all clustering based 
temporal summarization systems. In order to avoid this, we set a threshold $t$
such that a new update must have a maximum similarity $< t$ to all previous 
updates.

We determined the value of $t$ by examining the distribution of similarities
of nuggets to their human-matched sentences compared to the distribution
of similarities of nuggets to relevant but non matching sentences. We fit
normal distributions for the matching and non-matching similarities and
used the point of intersection as the value of $t$.
Figure ? shows a normalized histogram of cosine similarities for both 
distributions, the fitted distributions, and the point of intersection.

\begin{figure}
\includegraphics[scale=.40]{match-dist.png} 
\caption{Distribution of nugget similarity for matching and non-matching 
updates.\fdcomment{possible to output this as a vector graphic (pdf pref).  it will scale better.}}
\end{figure}
\subsubsection{Effect of Salience Models}

In order to assess the effectiveness of the salience predictions on update
selection, we perform a baseline (AP-base) run using affinity propagation 
clustering
with uniform preference, i.e. all preferences are equal to the median sentence
similarity. This run is compared to our full system (AP-sal), using each
sentence's salience prediction as its preference.

We are also interested
in the effect of directly incorporating salience in the clustering algorithm.
A third run (HAC) is performed using hierarchical agglomerative 
clustering; we produce flattened clusters by setting a maximum cluster 
distance threshold. The sentence with the highest predicted salience from each
 cluster is selected as an update. 
The distance threshold was tuned using grid search. By comparing
HAV with AP-base we can empirically assess the trade off between preference
and neighbor similarity that affinity propagation is making.

\subsubsection{Feature Ablation}

We remove each feature group and train models models on these small faller 
feature susbstets. All other parameters are the same as in our full system 
(AP-sal).

>>>>

In order to evaluate an update summarization system, we adopt the simulation-based approach used in the TREC Temporal Summarization (TS) Track.  We provide a brief overview of the problem.  Details on the formulation can be found in the track overview paper \cite{aslam2013trec}.

An update summarization system takes as input 
\begin{enumerate*}[label=\itshape\alph*\upshape)]
  \item a short query $\query$ defining the event to be tracked (e.g. `Hurricane Sandy'), 
  \item an event category $\eventcategory$ defining the type of event to be tracked (e.g. `hurricane'), 
  \item a stream of time-stamped documents, $(\doc_0, \doc_1,\ldots,\doc_T)$,
  presented in temporal order, and \item an evaluation time period of interest,
      $(\stime,\etime)$.  \end{enumerate*} While processing documents
      throughout the time period of interest, the system must output sentences,
      known as updates, which are \emph{relevant} to the query,
      \emph{comprehensive} with respect to the event, \emph{novel} to the user,
      and \emph{timely} with respect to when the update occurred.  Precisely
      how to measure these properties will be discussed in Section
      \ref{sec:methods}.  In our version of this problem, we assume that that
      the system receives one batch of new sentence-segmented documents every
      hour throughout the period of interest.

We address disaster types such as terrorism and mass shootings (e.g., the 2012
shooting in Aurora, Col.), natural disasters (e.g., Hurricane Sandy), accidents
(e.g., the 2012 Pakistan garment factory fire), and social activism (e.g., the
Arab spring). The size and scope of the events varies considerably. Hurricane
Sandy, for example, effected multiple countries over weeks while the
2012 Aurora shooting was contained to a single location spanning several hours.

%\kmcomment{This was as best as I could do on this. These are the example
%    categories from the wikipedia page. However you refer in the language model
%    description to the event type earthquake which is more specific. Also, in
%    the TREC data you refer to event types that are specified in the metadata.
%So I'm wondering how many different categories you have and where they come
%from.}

%\fdcomment{include description of nuggets here.} 
%%\kmcomment{We should jointly
%    discuss what goes in data and what goes in problem definition. I have moved
%    event types here because I think knowing how many types and examples of
%    them would be helpful upfront. I think where nuggets and wikipedia pages go
%is questionable.}

Figure \ref{alg:temporal-summarization} outlines our general update
summarization algorithm.  At each hour, the system processes each input 
sentence batch $S_t$.
We predict the salience $P$ for all input sentences $s\in S_t$.
Next, we cluster $S_t$ using the AP clustering algorithm, biased by $P$,
obtaining a set of exemeplar sentences $E$. Finally we select a subset of $E$
to be updates and add those to our set of summary updates $U$.

\begin{algorithm}%[H]
 \KwData{ 
     $S_{\stime}, S_{\stime+1},\ldots,S_{\etime} $ --- time ordered sentence
     batches\\
 \KwResult{U --- a list of updates, i.e. the update summary} 
}
 ~\\
 Initialize empty list $U$ of updates\\
    \For{$t \gets \stime,\ldots,\etime $}{
        $P \gets \operatorname{PredictSalience}(S_t)$\\
        $E \gets \operatorname{APCluster}(P, S_t)$ \\
        $U_t \gets \operatorname{SelectUpdates}(E,P,U)$\\
        $U \gets U \cup U_t$
}
 \caption{Temporal Summarization Algorithm}\label{alg:temporal-summarization}
\end{algorithm}

\subsection{Salience Prediction}

\subsubsection{Model}
%KM - I think one of the words ``sentence'' below should be something else, btu
%not sure what you meant. Ah.. I think ``salience''?
%We adopt Gaussian process regression in order to predict sentence sentence
We adopt Gaussian process regression in order to predict sentence salience
\cite{rasmussen:gaussian-process-book}.  Gaussian process regressors are a
class of data-driven, non-parametric model generalizing the multi-variate
Gaussian to the infinite dimensional setting.  Gaussian processes are general
and are state of the art for many regression tasks.  A full treatment of
Gaussian process regression is beyond the scope of this article and can be
found in standard textbooks.

As a non-parametric, the Gaussian process relies on a  covariance matrix $\kernelMatrix$, measuring the affinity between pairs of instances, in our case candidate sentences.  In our experiments, we used a radial basis function (RBF) kernel.  Given two featurized sentences, $\features$ and $\features^\prime$, the RBF kernel is defined as,
\begin{align*}
        \kernelMatrixij{\features}{\features^\prime}&= \sigma^2 \exp\left(- \frac{1}{2} 
\sum_{i=1}^{\numfeatures} \frac{ (\featuresi{i}-\featuresi{i}^\prime)^2}{\ell_i^2} \right)
\end{align*}
where $\sigma$ and the $\ell_i$ are parameters we fit to our observed training data. The $\ell_i$ are feature dependent scaling parameters; once learned, they not only improve the accuracy of the model, but give us some introspection  into which features are more important.

\subsubsection{Training Data}
\ckcomment{Explain how we train the model here...}
%In order to train the salience model, we need to quantify the notion of salience. We assume for each event we have a set of atomic facts, or nuggets.

%
\begin{figure}[t!]
\begin{tabular}{| l |} 
\hline
\textbf{Basic Features}\\
$\cdot$ Sent. position (normalized by doc. length) \\
$\cdot$ Sent. length \\
$\cdot$ Ratio of punc. to non-punc. chars. \\
$\cdot$ Ratio of caps. to non-caps. chars. \\
$\cdot$ Ratio of lowercase to other chars. \\
$\cdot$ Ratio of uppercase to other chars. \\
$\cdot$ \# of caps. words (normalized by \# of words)\\
$\cdot$ \# of Person, Location, Org. Date, Number,\\
$\;\;$ Ordinal, Percent, Money, Set, Misc  N.E. \\
$\;\;$ tags (normalized by \# of words)\\
\hline
\textbf{Query Features}\\
$\cdot$ \% of query words covered by sent.\\
$\cdot$ Total query matches.\\
$\cdot$ Total event-type synonyms/hypernyms/ \\
$\;\;$ hyponyms coverage.\\
$\cdot$ Total event-type synonyms/hypernyms/ \\
$\;\;$ hyponyms matches.\\
\hline
\textbf{Language Model Features}\\
$\cdot$ Avg. token log probability (domain lang. \\
$\;\;$ model)\\
$\cdot$ Avg. token log probability (background \\
$\;\;$ lang. model)\\
\hline
\textbf{Geo-tag Features}\\
$\cdot$ Median document distance to nearest  \\
$\;\;$ location cluster (current hour).\\
$\cdot$ Median document distance to nearest  \\
$\;\;$ location clusters (previous hour).\\
$\cdot$ Distance of first location in doc. to nearest \\
$\;\;$ location cluster (current hour).\\
$\cdot$ Distance of first location in doc. to nearest \\
$\;\;$ location cluster (previous hour).\\
\hline
\textbf{Temporal Features}\\
$\cdot$ Avg. tf-idf at current time.\\
$\cdot$ Change in avg. tf-idf since previous hour \\
$\;\;$  (up to 24 hours).\\
$\cdot$ Time since query/event start.\\
\hline
\end{tabular}
\caption{Salience model features.}
\end{figure}

\subsubsection{Features}
We want our model to be predictive across different kinds of events so we avoid lexical features.  Instead, we extract a variety of features including language model scores, geographic relevance, and temporal relevance from each sentence.  These features are used to fit a Gaussian process regression model that can predict the similarity of a sentence to a gold summary \cite{preotiuc2013temporal}.

\paragraph{Basic Features}
%KM - Would be good to have quick justification of these features. I added a
%sentence. Feel free to edit or remove.

We employ several basic features that have been used previously in supervised models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}. These include sentence length, the number of capitalized words normalized by sentence length, document position, number of named entities.  
Since training is one on grammatical English, some of these features help
to downweight sentences that are ungrammatical (e.g., have too many capitalized words or are too short).
Others help to more heavily weight important sentences (e.g., that appear in
prominent positions such as paragraph initial or article initial).

\paragraph{Query Features}

Query features measure the relationship between the sentence and the event query and type.  These include the number of query words present in the sentence in addition to the number of event type synonyms, hypernyms, and hyponyms as found in WordNet \cite{miller1995wordnet}.  For example, for event type \emph{earthquake},  we match sentence terms ``quake'', ``temblor'', ``seism'', and ``aftershock''.
\paragraph{Language Model Features}\label{subsubsec:lm}
Language models allow us to measure the likelihood of a sentence having been produced from a particular source.  We consider two types of language model features.  The first model is estimated from a corpus of generic news articles.  This model is intended to assess the general writing quality (grammaticality, word usage) of an input sentence and helps us to filter out text snippets which are not sentences (e.g., web page titles).  The second model is estimated from text specific to our event types.  For example, the language model for event type `earthquake' is estimated from Wikipedia pages under the category \emph{Category:Earthquakes}.  These models are intended to detect sentences similar to those appearing in summaries of other events in the same category (e.g. most earthquake summaries are likely to include higher probability for ngrams including the token `magnitude').  
%KM - Note: Someplace the exact list of event types should appear. Probably not
%here.
%KM - I note you have it in a later section but it is labeled as data you use
%to train language models and semantic similarity. I think it would be good to
%have up front in definition of task.

%For both models, we Finally, we extract the percentage of capitalized words,
%and sentence length as features. These last two features also help to
%identify sentences that are less likely to contain relevant content-- overly
%long and heavily capitalized sentences in our corpus were likely to be long
%strings of web-page headlines, section headers, and other irrelevant page
%structure.

\paragraph{Geographic Relevance Features}

Locations are identified using a named entity tagger. For each location in a sentence, we obtain its latitude and longitude using the a publicly available geolocation service.  We then compute its distance to that of the event location.  It is possible for a sentence and an event to have multiple locations so we take as features the minimum, maximum, and average distance of all sentence-event location pairs.  Distances are calculated using the Vincenty distance. 
%KM - Probably should say how you determine event location. Some events move.
%KM - In your figure you only include one distance.

\paragraph{Temporal Relevance Features}

Our data consists of hourly crawls of online content and so we exploit the temporality of corpus by capturing the burstiness of a sentence, i.e.  the change in word frequency from one hour to the next.``Bursty'' sentences often indicate new and important data.

Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,w_n\}$ be a sentence from a page $d \in D_t$.  We calculate the 1-hour burstiness of sentence $s$ from document $d$ at hour $t$  as 
%\begin{align*}
%\operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
%\operatorname{tf-idf}_t(w,d)  \\ & \left. - \frac{\sum_{d^\prime \in D_{t-1}:
%w \in d^\prime } \operatorname{tf-idf}_{t-1}(w,d^\prime)}{|\{d^\prime \in
%D_{t-1}: w \in d^\prime\}|} \right) \end{align*}

%where \begin{align*} \operatorname{tf-idf}_t(w,d) =&
%\log\left(1+\sum_{w^\prime \in d}1\{w=w^\prime\}  \right)\\ & \times
%\log\left(\frac{|D_t|}{1 + \sum_{d^\prime \in D_t}1\{w \in d^\prime\}}\right).
%\end{align*}

[\textrm{Simpler explanation goes here}]
% 1\{w = w^\prime} %- \operatorname{avg-tf-idf}_{t_{i-1}}(w).
%\end{align*}

We similarly find the sentence's 5-hour burstiness.  In addition to burstiness, we also include the sentence's average tf-idf and hours since the event in question started as features.

% Formally, let $p(f)$ be a distribution over functions where $f$ is any mapping
% of an input space $\mathcal{X}$ to the reals,
%
% $$f: \mathcal{X} \rightarrow \mathcal{R}.$$
% Let the random variable $\mathbf{f} = (f(x_1),\ldots,f(x_n) )$ be
%  an $n$-dimensional vector whose elements are evaluations of the function $f$
% at points $x_i \in \mathcal{X}$.
% We say $p(f)$ is a Gaussian process if for any finite subset
% $\{x_1,\ldots,x_n\} \subset \mathcal{X}$, the marginal distribution over
% that finite subset $p(\mathbf{f})$ has a multivariate Gaussian distribution.
% A GP is parameterized by a mean function $\mu(\mathbf{x})$ and a
% covariance function $K(x,x^\prime)$. Generally, the mean function is simply
% set to 0, leaving the distribution to be completely characterized by the
% kernel function on the data.
%
% In the regression setting, we typically have a response variable $y$ that
% is the sum of our model prediction  and
% some Gaussian noise, i.e. $y = f(x) + \epsilon$ with
% $\epsilon \sim \mathcal{N}(0, \sigma^2)$. When
% $f \sim \operatorname{GP}(\mathbf{0}, \mathbf{K})$, the
% two distributions
% of principal interest are the marginal likelihood
% $p(\mathbf{y}|\mathbf{X}) =
% \mathcal{N}(\mathbf{0},\mathbf{K} + \sigma^2\mathbf{I})$ and the predictive
% distribution,
%
% $$p(\mathbf{y_*}|\mathbf{x_*},\mathbf{X},\mathbf{y}) =
% \mathcal{N}(\boldsymbol{\mu}_*, \boldsymbol{\sigma}^2_*) $$
%
% where $\mathbf{x_*}$ is a new or unseen input, $\mathbf{y_*}$ our predicted
% response, and
% \begin{align*}
% \boldsymbol{\mu}_* & = \mathbf{K_*}(\mathbf{K} + \sigma^2\mathbf{I})^{-1}\mathbf{y} \\
% \boldsymbol{\sigma}^2_* &
% = \mathbf{K}_{**} - \mathbf{K}_*(\mathbf{K} + \sigma^2\mathbf{I})^{-1}
% \mathbf{K}_*^T + \sigma^2\\
% \end{align*}.
%
% Here $\mathbf{K}_* = K(\mathbf{x}_*, \mathbf{X})$, and
% $\mathbf{K}_{**} = K(\mathbf{x}_*, \mathbf{x}_*)$.
%
%

\subsection{Affinity Propagation Clustering}
\ckcomment{Explain AP and how salience gets plugged in here}
\subsection{Update Selection}
\ckcomment{Explain thresholds}
